A Dataset of Scratch Programs:\\Scraped, Shaped and Scored

### 2017-02-20 MSR-Scratch.tex
Some more minor fixes

While Scratch is receiving increasing interest as an introductory programming language, there is no recent dataset of Scratch programs available to the research community.
The one made available from the MIT Media Lab concerns projects created using the previous, initial version of the Scratch application, before the introduction of the current Scratch version (Scratch 2) and of the web programming interface in 2013.
It is since then that the popularity of Scratch started to increase\footnote{Monthly activity trends can be found at \url{https://scratch.mit.edu/statistics/}}. Scratch 2 introduced several features relating to important programming concepts,\footnote{The complete list of features introduced in Scratch 2 can be found at \url{https://wiki.scratch.mit.edu/wiki/Scratch\_2.0}} like custom blocks (the equivalent to procedure definitions), and therefore that dataset does not include projects utilizing those features.

@@

\subsection{Data collection}
To collect the data from the web interface of the Scratch project repository, we built a scraping program.
The web scraping program starts by reading the Scratch projects page\footnote{\label{scratchpublic}\url{https://scratch.mit.edu/explore/projects/all/}} and thus obtains the project identifiers of projects that were most recently shared.
Subsequently, it retrieves a JSON file for each of the listed projects\footnote{For a given project id $x$, the program's JSON representation can be obtained via \url{https://cdn.projects.scratch.mit.edu/internalapi/project/x/get}}.

@@

\subsection{Calculation of programming mastery scores}
%\fenia{Description of how the mastery metrics for       
%	[Abstraction]
%	,[Parallelism]
%	,[Logic]
%	,[Synchronization]
%	,[FlowControl]
%	,[UserInteractivity]
%	,[DataRepresentation]
%	,[Mastery]
%	,[Clones]
%	,[CustomBlocks]
%	,[InstancesSprites]
%	were calculated}
We analyzed the projects with Dr. Scratch, a tool that statically inspects Scratch projects' source code to assign scores on seven dimensions of computational thinking: abstraction and problem decomposition, logical thinking, synchronization, parallelism, algorithmic notions of flow control, user interactivity and data representation. These dimensions are given a value from 0 to 3. The mastery score is the aggregated punctuation of the seven dimensions, and therefore ranges from 0 to 21.

@@

\subsection{Importing the data}
\label{dataAnalysis}
All scraped project data and metadata, including the list of used blocks and parameters, were imported in a relational database.
We also imported the data on the shapes and the categories of the Scratch blocks.
We then used SQL queries for normalizing the data and bringing it in its final schema, outlined in Table~\ref{tbl-dbschema}.

>>>>

While Scratch is receiving increasing interest as an introductory programming language, there is no recent dataset of Scratch programs available to the research community.
The one made available from the MIT Media Lab concerns projects created using the previous, initial version of the Scratch application, before the introduction of the current Scratch version (Scratch 2) and of the web programming interface in 2013.
It is since then that the popularity of Scratch started to increase.\footnote{Monthly activity trends can be found at \url{https://scratch.mit.edu/statistics/}}
Scratch 2 introduced several features relating to important programming concepts,\footnote{The complete list of features introduced in Scratch 2 can be found at \url{https://wiki.scratch.mit.edu/wiki/Scratch_2.0}} like custom blocks (the equivalent to procedure definitions), and therefore that dataset does not include projects utilizing those features.

@@

\subsection{Data Collection}
To collect the data from the web interface of the Scratch project repository, we built a scraping program.
The web scraping program starts by reading the Scratch projects page\footnote{\label{scratchpublic}\url{https://scratch.mit.edu/explore/projects/all/}} and thus obtains the project identifiers of projects that were most recently shared.
Subsequently, it retrieves a JSON file for each of the listed projects\footnote{For a given project id $x$, the program's JSON representation can be obtained via \url{https://cdn.projects.scratch.mit.edu/internalapi/project/x/get}}.

@@

\subsection{Calculation of Programming Mastery Scores}
%\fenia{Description of how the mastery metrics for       
%	[Abstraction]
%	,[Parallelism]
%	,[Logic]
%	,[Synchronization]
%	,[FlowControl]
%	,[UserInteractivity]
%	,[DataRepresentation]
%	,[Mastery]
%	,[Clones]
%	,[CustomBlocks]
%	,[InstancesSprites]
%	were calculated}
We analyzed the projects with Dr. Scratch, a tool that statically inspects Scratch projects' source code to assign scores on seven dimensions of computational thinking: abstraction and problem decomposition, logical thinking, synchronization, parallelism, algorithmic notions of flow control, user interactivity and data representation. These dimensions are given a value from 0 to 3. The mastery score is the aggregated punctuation of the seven dimensions, and therefore ranges from 0 to 21.

@@

\subsection{Importing the Data}
\label{dataAnalysis}
All scraped project data and metadata, including the list of used blocks and parameters, were imported in a relational database.
We also imported the data on the shapes and the categories of the Scratch blocks.
We then used SQL queries for normalizing the data and bringing it in its final schema, outlined in Table~\ref{tbl-dbschema}.
